Tor's Usability for Censorship Circumvention

### 2015-12-09 paper/circumvention-experiment.tex
Write configuration, method, and design sections

\noindent {\bfseries Overview}
The Tor Browser provides a configuration interface to guide users through setting up 
their connection to the
Tor network. While people in most countries can connect to Tor using the
dialog's default settings, citizens in certain censorship environments must
make use of additional settings, such as bridges, proxies, or both. 
{\color {red} definition of a bridge, definition of a proxy}   %%%
Our study  will explore how well this interface successfully guides users to 
correctly configuring their browser.\\

@@

\noindent {\bfseries Configuration Flow}
There are five general paths a user can take through the Tor Browser connection
configuration interface. {\color {red} These paths are shown in Figure ~\ref{fig:interface}}. 
The configuration interface asks a series of questions to 
the user to determine if the user requires a bridge or proxy when connecting to 
the Tor network. Depending on how the user answers these questions, a user 
has give paths through the configuration, as follows:

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item {\bfseries Path 1}:  Connect directly
\item {\bfseries Path 2}:  Yes bridge, no proxy, connect
\item {\bfseries Path 3}:  Yes bridge, yes proxy, connect 
\item {\bfseries Path 4}:  No bridge, yes proxy, connect 
\item {\bfseries Path 5}:  No bridge, no proxy, connect 
\end{itemize}

\begin{figure*}[t]
\label{fig:interface}
  \centering
    \includegraphics[width=\textwidth]{../torconfig.png}
    \caption{This flow chart shows all of the possible paths taken through the
    Tor configuration interface. A state represents a window in the interface,
    with the exception of the leaves, which indicate the final action taken by
    the interface (``q'' for quit, ``s'' for success, and ``e'' for error). The
    transitions between the states indicate which user action causes that
    transition. The cause for error is can be found
\href{https://github.com/lindanlee/circumvention-ux-tor/blob/master/torconfig.dot}{here}.}
\end{figure*}

Upon the startup of the configuration interface, a user is asked to describe their 
current situation. If their Internet connection is censored or proxied, the user 
clicks ``Configure'' to configure bridge and proxy settings. Otherwise, the user
clicks ``Connect'' to connect directly to the Tor network.

If a user is not connecting directly to the Tor network, then they are asked to 
configure a bridge (as seen in {\color {red} TODO}, the title of the window during configuration %%%
is ``Tor Bridges Configuration''). We suspect that the reason for configuring a bridge
first before a proxy is that more people need to configure bridges than proxies,
but this is not known. The bridge configuration asks the users if their ISP blocks
or censors connections to the Tor Network. The assumption here is that users will 
not know if they need a bridge or not, but that they are equipped to answer if their
current network environment censors the Tor Network. The configuration interface 
uses this answer to deem if the user needs a bridge. A yes to this question would
require a bridge while a no to this question would not. If they answer yes, they are 
asked to configure the bridges. If they answer no, then the configuration interface 
moves on to configuring proxies.

If a user who is not directly connecting to the Tor network is finished configuring
a bridge or has answered that they do not need a bridge, they are asked to 
configure a local proxy (as seen in {\color {red} TODO}, the title of the window during %%%
configuration is ``Local Proxy Configuration''). The proxy configuration asks the 
users if they need to use a local proxy to access the Internet. The assumption 
is that people will not know the answer to this question, so there is a text instructing
users how to find the answer to this question below the question. If the answer to
the question is yes, the user needs a proxy and is taken to a screen where they are
asked to configure the proxy settings. If the user does not need a proxy, then the 
configuration interface attempts to connect to the Tor network with all the 
given inputs up to this point so far.

Note that a user can decide to click ``Configure'' because they believe the 
Internet connection is censored or proxied, but then answer the bridge and proxy
questions in such a way that they deemed that they do not need a bridge or proxy 
after all. In this case, clicking the ``Connect'' button after the above configuration steps
would be equivalent to clicking ``Connect'' on the first screen. \\

\noindent {\bfseries Things that are good about the interface} %% fix this so it's not so casual
{\color {red} 
Talking to users, ``this will work in most situations,'' help button, support email.
putting the bridge configuration before the proxy, etc.... \\
}

\noindent {\bfseries Things that could be improved about the interface }%% fix this so it's not so casual
{\color {red} 
clarify between bridges and proxy, getting rid of redundant text, etc.  \\
}

\noindent {\bfseries Hypotheses} We expect that the average user will
struggle to connect because they lack knowledge of the technical terms and
mechanisms required for circumvention (such as pluggable transports) and cannot
obtain information not explicitly provided by the user interface (e.g., IP
addresses of non-publicly listed bridges).

@@

We are going to aim for 5 users per censorship environment, improvement, 
and alternate interface. We are going to pre-screen our participants so that 
we have a good mix of gender, age, technical background, and familiarity 
with Tor.

Some references to keep in mind when explaining: 
\href{why 5 users}{http://www.nngroup.com/articles/how-many-test-users/}
\href{screening}{http://www.userfocus.co.uk/articles/screeners.html}
\href{summative and usability mixed model}{http://www.usabilitybok.org/summative-usability-testing}

\noindent {\bfseries Experiment Logistics} 
The IRB protocol to run this user study has been approved (2014-12-6995). 
We plan to recruit 35-50 users for the purpose of this study,
making it the largest user study of Tor to date.  This study will be conducted at the
Experimental Social Science Laboratory (Xlab)
at the University of California, Berkeley, which consists of 36 laptops,
separated by cubicle walls. We will use individual host firewalls to simulate
censorship environments and will record computer screens to capture 
user activity. The total length of the experiment, including briefing, completing the censorship 
circumvention tasks, exit survey, and debriefing, will be about an hour.
Participants will be compensated \$30 for their time, which covers
minimum wage for an hour and any transportation costs to the lab.  \\

\noindent {\bfseries Experiment Flow} 
We will conduct three separate experiments as part of this study:
the first set to test the current configuration
interface, a second set to test the improved interface, and a third set to test the 
alternate configuration interface. The experiment flow will be the same across all 
sets of experiments, but the configuration interface will be different across sets.

The experiment begins with the participants being informed that they are in a
simulated censorship
environment, where some --- but not all --- websites are blocked. We will
instruct them to visit a non-blocked website and a blocked website on a
``standard'' browser (one that is not used for censorship circumvention, such
as Chrome, Safari, or Firefox) to illustrate the situation.
Then, the participants will find out the role they they will be playing in the study.
This role is that of a regular citizen who is trying to visit blocked websites
for non-critical leisurely purposes,
and for whom the chance of risk is minimal. This information is given
to achieve
consistency across participants: behaviors may change if the tasks were 
critical or if there were great risks associated with configuration mistakes.

After explaining the situation and their role, we will explain what Tor is, and
how they can use it if necessary, and instruct them to complete a set of
tasks which will require visiting blocked websites. Ultimately, the
participants will need to configure their Tor Browser to circumvent the
simulated censorship environment. These tasks have a three-fold purpose of
providing participants with feedback (if they did not correctly configure their
browser, they cannot reach the website),  motivating the participants to
configure the browser correctly, and providing the researchers with an
indicator of successful censorship circumvention.
Participants' screens will be recorded to capture how they configure
their browser, and to provide additional evidence that they have completed the
tasks.

After users complete the browsing tasks, we will administer a survey that 1)
measures demographics, technical fluency, and
familiarity with Tor and 2) collects qualitative feedback from
participants about their browsing experience as a whole
and the configuration interface specifically.
A rough draft of the survey can be
found
\href{http://www.surveygizmo.com/collab/2085559/Tor-Usability-Survey}{here}.

The experiment will conclude with a debriefing, which will inform that the
participants of their screen capture and obtaining re-consent for the
information. \\

\noindent {\bfseries Censorship Environment Simulation} 
We plan to simulate three censorship environments.
They are informed by our experience with pluggable transports
and knowledge of commonly seen censorship techniques.
They are not meant to perfectly replicate the network environment
in any particular country. Although inspired by reality, these
abstract simulations are intended to require distinct configurations
of the Tor Browser from our participants. \\

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item {\bfseries Mild censorship} 
(Reflective of countries such as France and Australia.)
Certain domains are blocked. Reaching these 
domains requires a censorship circumvention 
tool. The default option to ``connect'' to the Tor network 
directly will circumvent this censor. Additional correct
bridge or proxy configurations are optional. 
\item {\bfseries Intermediate censorship} 
(Reflective of countries such as Tunsia.)
Certain domains are blocked. Censorship circumvention
tools are blocked. Since Tor is blocked by blocking all public Tor
relay nodes, the default option to ``connect'' to the Tor network
directly will fail. Any choice of a hard-coded bridge (see Figure ~\ref{fig:bridges})
or a valid non-public bridge will circumvent this censor.  
Additional correct proxy configuration is optional.
\item {\bfseries Comprehensive censorship} 
(Reflective of countries such as China and Syria.)
Certain domains are blocked. Censorship circumvention tools
are thoroughly blocked. Tor is blocked by blocking all public
Tor relay nodes, and the censor has examined source code to block
all hard-coded bridge relays in the configuration interface. The default option
to ``connect'' to the Tor network directly will fail. Choosing any bridges other than
``meek-amazon,'' ``meek-azure,'' and ``meek-google,'' will fail. This is because 
domain-fronting requires censors to block entire CDNs to also block this
transport (which will cause huge collateral blocking damage), making it resistant to agressive censorship environments.
(See ~\cite{fifield2015blocking} for additional details.)\\
\end{itemize}

\noindent {\bfseries List of Tasks} 
In our experimental setup, successful completion of 
these given tasks requires of correct configuration.
The difficulty of circumventing  censorship to visit the 
websites required for the tasks will vary depending on the 
simulated censorship environment. Since all participants will be given the 
same set of tasks, regardless of the censorship environment, 
the difficulty of completing the tasks remains constant assuming
correct configuration. The tasks themselves are not intended to
be challenging.

The tasks were inspired by the top Alexa sites,
an indication of representative and relevant browsing behavior. 
From these, we selected sites which were commonly censored, 
but filtered tasks that would require a participant to reveal private information 
(such as login information) for ethical considerations. These tasks
were further refined after performing a pilot study of the experiment. 
We hope the tasks convey an example of how of a user might 
browse the Internet in a censored environment.

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item Google search for the population of Zimbabwe. 
\item On YouTube, find a video playing Bach's ``Ode to Joy.''
\item Find the Amazon best-sellers in ``Movies \& TV.''
\item On Yahoo, find the exchange rate of dollars to euros.
\item Find the Wikipedia ``History'' portal's featured article. 
\item On Twitter, find the currently trending topics.
\item On Bing Maps, find directions from Times Square to Carnegie Hall.\\
\end{itemize}

\noindent {\bfseries Improved and Alternative Interfaces} We have yet to design the 
improvements to the interface, as they will be based on the data collected from 
the user studies on the current configuration interface. In general, we will measure the 
time it took to configure correctly, and help users avoid common mistakes. An alternate 
interface to automate the configuration process for the user has been proposed in the 
community. One concern about it is that the automation makes the connection process less transparent, 
and since there are countries who log Internet activity, automated attempts to obviously
use the Tor network could have the potential to be incriminating.\\

\noindent {\bfseries Exit Survey}
{\color {red} talk about the survey here.}

@@

>>>>

\noindent {\bfseries Overview}
The Tor Browser provides a configuration interface to guide users through setting up 
their connection to the
Tor network. While people in most countries can connect to Tor using the
dialog's default settings, citizens in certain censorship environments must
make use of additional settings, such as bridges, proxies, or both. 
(Bridges are unlisted Tor relays, with pluggable transports to obfuscate traffic.
Proxies help bypass local network restrictions and are typically only
needed in restricted cases such as corporate environments.)
Our study explores how successfully this interface guides users to
correctly configuring their browser.\\

\noindent {\bfseries Configuration Flow}

\begin{figure*}[t]
\label{fig:interface}
  \centering
    \includegraphics[width=\textwidth]{../torconfig.png}
    \caption{This flow chart shows all of the possible paths taken through the
    Tor configuration interface. A state represents a window in the interface,
    with the exception of the leaves, which indicate the final action taken by
    the interface (``q'' for quit, ``s'' for success, and ``e'' for error). The
    transitions between the states indicate which user action causes that
    transition. Annotations for the error states can be found by following
\href{https://github.com/lindanlee/circumvention-ux-tor/blob/master/torconfig.dot}{this hyperlink}.}
\end{figure*}

As bridges and proxies can be configured independently or together, there are
four main ways to set up a connection:

\begin{enumerate} \itemsep1pt \parskip0pt \parsep0pt
    \item With a bridge and proxy
    \item With a bridge but no proxy
    \item Without a bridge but with a proxy
    \item Without either a bridge or proxy
\end{enumerate}

The launcher interface asks a series of questions to
determine which of these configuration states a user requires.
The four high-level states correspond to five distinct paths through the
configuration interface. (The latter option -- connecting with neither a bridge
nor proxy -- contributes two separate paths.) Each path branches further
depending on user-specific inputs. Figure \ref{fig:interface} illustrates the
resulting paths.

When the configuration window opens, a user is given the option to connect
directly to the Tor network by clicking ``Connect'', or to click ``Configure''
``if this computer's Internet connection is censored or proxied'' (Figure
\ref{fig:window1}).

If they choose the latter, the next screen asks, ``Does your Internet Service
Provider (ISP) block or otherwise censor connection to the Tor network?''
(Figure \ref{fig:window2}). Though this sounds repetitive, it is meant to
identify and divert users who need only a proxy (and no bridge).
This phrasing of the question carries the assumption that people will
not know if they need a bridge or not, but that they are equipped to answer if their
current network environment censors the Tor Network.

@@

Next, users are asked to configure bridges by selecting a transport type from a
dropdown menu or entering the IP address of a custom bridge in a textbox (Figure
\ref{fig:bridges}).
Finally, users are asked whether they need a local proxy (Figure
\ref{fig:window4}) and (if they answered yes) to configure it (Figure
\ref{fig:window5}).

Note that a user can decide to click ``Configure'' on the first screen, but then
opt out of both bridges and proxies.
In this case, connecting after these configuration steps
would be equivalent to clicking ``Connect'' on the first screen. \\

\noindent {\bfseries Design choices}
The launcher interface utilizes a number of techniques to aid users and simplify
the choices they have to make. Help buttons and a support email address that
appears on every screen are meant to assist users that get stuck in the process.
The interface also makes good use of defaults, for example by emphasizing that
certain options ``will work in most situations.''

Still, there remains potential for confusion, especially with the more advanced
configuration options. Bridges and proxies are not clearly defined, the
interface doesn't specify when they would need to be used, and there is limited
guidance for how to correctly choose specific values (e.g., which transport type
to use or where to find proxy settings).

\section{Hypotheses}
We expect that the average user will
struggle to connect because they lack knowledge of the technical terms and
mechanisms required for circumvention (such as pluggable transports) and cannot
obtain information not explicitly provided by the user interface (e.g., IP
addresses of non-publicly listed bridges).

@@

To test our hypotheses, we simulated several environments in which people may
have to use Tor for censorship circumvention. We examined whether users are able
to join the network and studied how they interact with the launcher interface.
Based on our findings, we designed improvements to the Tor launcher interface,
which we test in a large-scale study.

\subsection{Simulated censorship environments}
We simulated three censorship environments.
They are informed by our experience with pluggable transports
and knowledge of commonly seen censorship techniques.
They are not meant to perfectly replicate the network environment
in any particular country. Although inspired by reality, these
abstract simulations are intended to require distinct configurations
of the Tor Browser from our participants.

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item {\bfseries Mild censorship} 
(Representative of countries such as France and Australia.)
Certain domains are blocked. Reaching these 
domains requires a censorship circumvention 
tool. The default option to ``connect'' to the Tor network 
directly will circumvent this censor. Additional correct
bridge or proxy configurations are optional.

\item {\bfseries Intermediate censorship} 
(Representative of countries such as Tunsia.)
Certain domains are blocked. Censorship circumvention
tools such as Tor are blocked. Since all public Tor
relay nodes are blocked, the default option to ``connect'' to the Tor network
directly will fail. Any choice of a hard-coded bridge
or a valid non-public bridge will circumvent this censor.  
Additional correct proxy configuration is optional.

\item {\bfseries Comprehensive censorship} 
(Representative of countries such as China and Syria.)
Certain domains are blocked. Censorship circumvention tools
are thoroughly blocked. Tor is blocked by blocking all public
Tor relay nodes, and the censor has examined source code to block
all hard-coded bridge relays in the configuration interface. The default option
to ``connect'' to the Tor network directly will fail. Most bridges will fail,
but ``meek-amazon,'' ``meek-azure,'' and ``meek-google'' still work.
% TODO: meek-google doesn't work either, right? Should we remove it?
This is because domain-fronting requires censors to block entire CDNs to also
block this transport (which will cause huge collateral blocking damage), making
it resistant to aggressive censorship environments.
(See ~\cite{fifield2015blocking} for additional details.)\\
\end{itemize}

\subsection{Qualitative interface evaluation}
To identify problems with the existing interface, we conducted a series of
qualitative evaluations with a set of users representative of the general
population.

Using established best practices from the field of user experience research
(\cite{howmanyusers}), we recruited five users for each censorship environment.
We pre-screened our participants to have a good mix of gender, age, technical
background, and familiarity with Tor. \\

%Some references to keep in mind when explaining:
%\href{why 5 users}{http://www.nngroup.com/articles/how-many-test-users/}
%\href{screening}{http://www.userfocus.co.uk/articles/screeners.html}
%\href{summative and usability mixed model}{http://www.usabilitybok.org/summative-usability-testing}

\noindent {\bfseries Procedure}
The experiment begins with the participants being informed that they are in a
simulated censorship
environment, where some --- but not all --- websites are blocked. We
instruct them to visit a non-blocked website and a blocked website on a
``standard'' browser (one that is not used for censorship circumvention, such
as Chrome, Safari, or Firefox) to illustrate the situation.
Participants are told that their
role is that of a regular citizen who is trying to visit blocked websites
for non-critical purposes,
and for whom the chance of risk is minimal. This information is given
to achieve
consistency across participants: behaviors may change if the tasks are
critical or if there were great risks associated with configuration mistakes.

After explaining the situation and the user's role, we
instruct them to complete a set of
tasks which will require visiting blocked websites. Ultimately, the
participants will need to configure their Tor Browser to circumvent the
simulated censorship environment. These tasks have a three-fold purpose of
providing participants with feedback (if they did not correctly configure their
browser, they cannot reach the website), motivating the participants to
configure the browser correctly, and providing the researchers with an
indicator of successful censorship circumvention.
Participants' screens are recorded to capture how they configure
their browser.

After users complete the browsing tasks (or time runs out), we conduct an
interview, collecting qualitative feedback from
participants about their browsing experience as a whole
and the configuration interface specifically.
We ask questions such as
``What was the most challenging part of connecting?''
and
``What is one change you would recommend?''

\subsection{Interface design and iteration}
After completing the qualitative user studies, we analyzed their results to
identify common themes and problems. Based on these, we then designed
improvements to the Tor launcher interface.

To do this, we used an iterative design process, starting with low-fidelity
(paper and pencil) prototypes and then proceeding to higher-fidelity versions.
At each stage, the prototype was clickable, allowing for more realistic
simulation.

We collected feedback from researchers, professors, and Tor users and developers
through the Tor project's mailing lists. We then iterated on the design by
incorporating the comments and suggestions we received into the next version of
the interface.

\subsection{Quantitative evaluation of improvements}
Once the new design of the user interface is finalized, we will conduct a
large-scale quantitative evaluation, to test whether the improvements had a
measurable effect on people's ability to connect to the Tor network.

The experimental procedure will be nearly identical to the procedure for the
qualitative portion of the study, described above. However, rather than
conducting open-ended interviews at the end of the tasks, participants will be
asked to fill out a brief survey. Furthermore, the interface will be
instrumented to collect metrics and statistics for use in analysis and
hypothesis testing.
Some examples of metrics collected include the time spent on each screen, which
paths the user followed, which interface elements they interacted with, and how
often they did so. These will be analyzed in conjunction with high-level metrics
(was the user able to connect to Tor successfully?).

Participants will be split across several different experimental conditions.
Some will be testing the improved user interface, while those in the control
group will be completing the same tasks using the old, unimproved interface. \\
% NOTE: removed mention of the alternate interface, unless there's something to say about it

\noindent {\bfseries Experiment Logistics} 
We plan to recruit 35-50 users for the purpose of this study,
making it the largest user study of Tor to date.  This study will be conducted at the
Experimental Social Science Laboratory (Xlab)
at the University of California, Berkeley, which consists of 36 laptops,
separated by cubicle walls. We will use individual host firewalls to simulate
censorship environments and will record computer screens to capture 
user activity. The total length of the experiment, including briefing, completing the censorship 
circumvention tasks, exit survey, and debriefing, will be about an hour.
Participants will be compensated \$30 for their time, which covers
minimum wage for an hour and any transportation costs to the lab. 
The IRB protocol to run this user study has been approved (2014-12-6995).

@@

\section{Design}
Based on the feedback we received and the insights we gathered from our
participants, we designed improvements to the launcher interface. This section
discusses our results and the design considerations we employed. \\

\noindent {\bfseries Automation}
One common piece of feedback, from experts and novices alike, was that the
connection process could be simplified with automation. At first glance, many
decisions currently made by the user could be automated. For example, bridge
selection needs to happen only if a regular connection failed, and bridges could
be tried automatically until one succeeds. However, there are security
considerations that make this solution problematic. If the launcher starts by
trying to make unobscured connections to publicly listed relays, detection is
trivial in a country that logs Internet traffic, potentially resulting in
repercussions for citizens. Furthermore, some users may be turned off by the
lack of transparency in the connection process.

However, other elements of the process can be automated without adverse effects.
For example, local proxy settings can be imported from other browsers on the
machine by looking for settings in well-known locations on the filesystem. The
same can be done to answer the question that led many of our participants
astray: is a local proxy even needed? Since this piece of automation offers a
significant potential upside, with no apparent risks, we have incorporated it
into our design. \\

\noindent {\bfseries Building a better mental model}
Participants' lack of a mental model of censorship led them to believe their
connection was censored when, in reality, it was not. It furthermore contributed
to their confusion between bridges and proxies. Our design directly addresses
these issues.

Potentially ambiguous text (i.e., referring to censorship) is removed in favor
of stating the underlying options. Separate screens with binary options (Do you
need a proxy? Yes/No) are removed; instead the option to opt out is incorporated
on the configuration screen. This results in a more streamlined interface (fewer
clicks required) and unified the context of each option: all information about
bridges (or proxies) is now contained on one screen, so the user has the full
context of their decision. \\

\noindent {\bfseries Clarifying paths through the interface}
Many users struggled because the workflow implicitly suggested by the interface
didn't match the ideal path: crucial steps to follow were unclear or obscured,
while undue prominence was given to comparatively less important elements.
Our new design inverts this trend, focusing attention on the interface elements
that are more important, or more generally applicable.

A clear example of this is the ``Connect'' button on the first screen, which has
increased in size and become green, making it more prominent as the recommended
option. A similar aesthetic applies to the connect button seen after custom
configuration.

Rarely used elements are deemphasized. For example, the custom bridges box,
which previously took up a significant portion of the bridge configuration
screen, is now hidden until that option is selected.

A progress indicator at the top of the window indicates which configuration step
the user is on. It serves to clarify the independence of bridges and proxies,
which need to be configured separately. In the event of a failure, the user is
returned to the first step of the process -- the bridge configuration -- to
avoid misleading users into thinking the proxy settings are the culprit. \\

\noindent {\bfseries Designing for failure cases}

Many users were led astray by interface elements that were unhelpful or
misleading if the program encountered connection problems. Our design seeks to
prevent such confusion by gracefully handling failures.

The progress bar in the connection window drew many complaints, because it got
stuck and didn't offer any guidance for how long users were supposed to wait. We
address this by providing a countdown timer, indicating how long users should
expect to wait. We further enhance the progress bar by adding visual indicators
for major milestones in the connection, providing better feedback about what has
been accomplished so far. Additionally, the current status of the connection
attempt (drawn from the log) is displayed, to aid in understanding and
debugging.

In the event that the connection attempt fails, messages from the log are
displayed, along with tips and suggestions about what to do next (e.g., try
another bridge). To make potential the options clearer, the new failure screen
includes a separate button for retrying without changing settings and another
for returning to reconfigure the application. \\

\noindent {\bfseries Additional changes}
In addition to the changes above, we have simplified flows where possible,
provided clearer definitions of bridges and proxies, and reduced visual clutter.
{\color {red} Anything else?}
